\setcounter{page}{1}
\thispagestyle{plain}
\begin{center}
	\vspace*{\fill}
	\textbf{Acknowledgements}\\
\end{center}
	I would like to express my gratitude to my advisor, Professor Enrique Peacock-L\'opez for supporting me throughout my thesis. It was always helpful to have someone I could bounce ideas off of and to explain to me obtuse math I would find in the literature. You really let me develop and understand the material needed independently but were always there when I needed help. You were the first professor I did research with in the summer of my freshman year and it was a pleasure to work on my thesis with you in my senior year.
	
	I would also like to thank my suite and the people of the Jesup community for providing me with both the emotional and technical support throughout my thesis. You provided the breaks after long hours of work and I look forward to staying close to all of you long after I graduate

	Thank you to the faculty and students Chemistry Department of Williams College for being with me all of these years. I understand that my thesis appears unorthodox for the department but I thank you all for being understanding and willing to let me work on this. Special thanks to Brandon Vuong, Si Hou Lon, and Alison Wong for making sure I understood what I could expect the average chemistry student to know and understand. 

	I of course must mention my family who were instrumental in raising me and helping reach the point I am now. Thank you to my mother and father, to my brother and sister: Helen and Bryant, and to the rest of my extended family to whom I love dearly.
	\vspace*{\fill}

\pagebreak